Dish: Distribution-oblivious Shell Scripting

### 2019-11-23 pldi/draft.tex
More on eval

\usepackage{tikz}
\usetikzlibrary{arrows,automata,shapes.misc,shapes.geometric,positioning}

@@

\newcommand{\kstar}{^{\textstyle *}}
\newcommand{\eps}{\varepsilon}

@@

For a set $D$, we write $D\kstar$ to denote the set of all finite
words over $D$. For words $x, y \in D\kstar$, we write $x \cdot y$
or $xy$ to denote their concatenation. We write $\eps$ for the empty
word. We say that $x$ is a \emph{prefix} of $y$, and we write $x \leq y$, if there is a word $z$ such that $y = xz$. It is easy to see that $\leq$ is a partial order, and it is often referred to as the \emph{prefix order}.

A node $f$ of the graph represents a function from one (possibly empty)
input stream to an output stream $f : D\kstar \to D\kstar$, where $D$
represents the basic data type of a line of characters. This
representation captures the majority of shell commands. We require that
the function $f$ is monotone w.r.t.\ the prefix order. This captures
the idea that a node cannot retract output after it has been emitted.
%More formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there exists $y'$ such that $f(x.x') = y.y'$, where $.$ is standard
%concatenation and will be omitted when obvious.
An example of a node is the following command:

@@

Edges in the dataflow graph represent files, the basic data
abstraction of the shell. They are used as communication channels
between nodes in the graph, and as the input or output of the whole
graph. Edges are represented as possibly unbounded streams of type
$D\kstar$. They could either refer to a specific named file in the file
system, or just be FIFO pipes that used for interprocess communication
%% \TODO{Are there any other file types, such as URLs, ...?}.

@@

\begin{center}
\small
\begin{tikzpicture}[node distance=2.25cm, ->]
\node (S) {};
\node (A) [draw, right of=S, node distance=1.75cm] {node 1};
\node (B) [draw, right of=A] {node 2};
\node (C) [draw, right of=B] {node 3};
\node (E) [right of=C, node distance=1.75cm] {};
\path (S) edge node[above] {input} (A);
\path (A) edge node[above] {stream} (B);
\path (B) edge node[above] {stream} (C);
\path (C) edge node[above] {output} (E);
\end{tikzpicture}
\end{center}

@@

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. This is the basic optimization
that is used by popular distributed processing systems such as
MapReduce \cite{DG2004MR} and Apache Spark\cite{spark:12}. Unfortunately, the
dataflow graph model of most of these systems is not expressive enough
to represent data parallelism \km{not sure this is true}, so the distributed implementations that
they perform do not preserve the semantics of their dataflow model.

@@

Experiments were run on a \todo{network of five workstations} connected by 1Gbps links: 
 one large machine (\wka) with 512GB of memory and 128 2.1GHz Intel Xeon E5-2683 cores, and 
 \todo{four smaller machines (\wkq), each with 4GB of memory and two 3.33GHz Intel Core Duo E8600 processors}.
For our software setup, we use SMP Debian 4.9.144-3.1 (2019-02-19), GNU Core-utils 8.30-3, GNU Bash 5.0.3(1)-release (x86\_64-pc-linux-gnu), Python 3.7.3, and OCaml 4.05.0.
Generally, no special configuration was made in hardware or software beyond disabling hyper-threading. % ---specifically, the network protocol stack was left unoptimized.
More detailed setup and experiments, including the scripts, is included in the accompanying (anonymized) online repository.

@@

\begin{table}[t]
\center
\footnotesize
% \setlength\tabcolsep{3pt}
\caption{
  \footnotesize{
    \textbf{Summary of Micro-benchmarks}.
    Six micro-benchmarks used to evaluate \sys.
  }
}
\begin{tabular*}{\columnwidth}{l @{\extracolsep{\fill}} lll}
\toprule
  Script                 ~&~ Structure                  & Input Size    & Seq.  Time \\
\midrule
  \tti{min-grep}         ~&~$2\times$\tsta              &               &            \\
  \tti{min-sort}         ~&~~\tpur                      &               &            \\
  \tti{top-n}            ~&~~\tpur                      &               &            \\
  \tti{wf}               ~&~~\tpur                      &               &            \\
  \tti{spell}            ~&~~\tpur                      &               &            \\
  \tti{diff}             ~&~~\tpur                      &               &            \\
\bottomrule
\end{tabular*}
\label{tab:eval}
\end{table}

We first apply \sys onto a collection of micro-benchmarks

Spell program~\cite{bentley1985spelling}

\ttt{wf}
Word-count~\cite{bentley1986literate}

\ttt{top-n}

The first benchmarks runs a grep over a series of complex, DFA
(Fig.~\cite{})

@@

\km{Some very early references on the dataflow model of computation (maybe there are relevant?): \cite{KM1966, D1974Dataflow, K1974KPN, KMacQ1977}}

\cite{HSSGG2014, SHGW2015} discuss optimizing transformations, but their correctness is established using informal arguments based on operational intution.

The paper \cite{MSAIT2019} proposes a denotational semantic framework for stream processing where the data streams are seen as partial orders, and establishes the soundness of some common parallelizing transformations on dataflow graphs.

>>>>

@@

@@

Nodes $f$ of the graph represent functions from one (possibly empty)
input stream to an output stream $f : D^* \rightarrow D^*$. $D$
represents the basic data type of a line of characters. This
representation captures the majority of shell commands. The nodes are
monotonic, meaning that they cannot retract their output. More
formally, $\forall f, x, y, f(x) = y$, then for any $x'$, there exists
$y'$ such that $f(x.x') = y.y'$, where $.$ is standard concatenation
and will be omitted when obvious. An example of a node is the
following command:

@@

Edges in the dataflow graph represent files, the basic data
abstraction of the shell. They are used as communication channels
between nodes in the graph, and as the input or output of the whole
graph. Edges are represented as possibly unbounded streams of type
$D*$. They could either refer to a specific named file in the file
system, or just be FIFO pipes that used for interprocess communication
%% \TODO{Are there any other file types, such as URLs, ...?}.

@@

@@

In order to adequately utilize the available computational resources,
a distributed implementation should exploit data parallelism. Data
parallelism is based on the observation that some operations produce
the same result when performed independently on subsets of the input
data, merging the results afterwards. This is the basic optimization
that is used by popular distributed processing systems such as
MapReduce\kk{cite} and Apache Spark\kk{cite}.Unfortunately, the
dataflow graph model of most of these systems is not expressive enough
to represent data parallelism, so the distributed implementations that
they perform do not preserve the semantics of their dataflow model.

@@

Experiments were run on a %\todo{network of five workstations} connected by 1Gbps links: 
 machine with 512GB of memory and 128 2.1GHz Intel Xeon E5-2683 cores.
%\todo{four smaller machines (\wkq), each with 4GB of memory and two 3.33GHz Intel Core Duo E8600 processors}.
For our software setup, we use SMP Debian 4.9.144-3.1 (2019-02-19), GNU Core-utils 8.30-3, GNU Bash 5.0.3(1)-release (x86\_64-pc-linux-gnu), Python 3.7.3, and OCaml 4.05.0.
Generally, no special configuration was made in hardware or software beyond disabling hyper-threading. % ---specifically, the network protocol stack was left unoptimized.
All pipelines are set to (initially) read from and (finally) write to the file-system (except as otherwise stated).
Whenever \ttt{curl} is part of a pipeline, it fetches data from a different physical host on the same network connected by 1Gbps links.
More detailed setup and experiments, including the scripts, is included in the accompanying (anonymized) online repository.

@@

Tab.~\ref{tab:eval} summarizes the collection of micro-benchmarks used to evaluate \sys.
Benchmark names (col. 1) correspond to the names on the plots (Fig.~\ref{fig:microbenchmarks}).
% The centralized pipelines read files of various sizes from disk and write to disk.

Input sizes and times (col. 2, 3) summarize the longest sequential execution (200 files in Fig.~\ref{fig:microbenchmarks}).
Script sizes report on the \sys's rewritten output for three different distribution sizes---$1\times$, 10$\times$, and $100\times$.

The size of the script grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
The first number alone is interesting, as it captures \sys's output for the sequential script.
As there is no parallelism, it can be viewed as the initial cost corresponding to using \sys.
% this is the size of 
This change in size does not necessarily translate to observed runtime, as most of this size gain comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
FIFOs are simply pipes made, the in-kernel buffering and synchronization mechanisms are the same, leading to non-observable changes in runtime performance.

In terms of benchmarks, \ttt{grep} is a short script centered around an expensive \sta command.
% a DFA-based regular expression matching phase;
  while \ttt{grep} defaults to a Thompson NFA, this particular expression makes use of DFA-based backtracking patterns resulting in high runtime overhead.
The \ttt{sort} pipeline is equally short, but its central operation is a command in \pur.
The next two benchmarks, \ttt{wf} and \ttt{top-n}, are based on McIlroy's now classic word-counting program~\cite{bentley1986literate};
  they use sorting, rather than tabulation, to identify high-frequency terms in a corpus.
The \ttt{bi-gram} pipeline calculates n-grams of base two;
  it makes clever use of \ttt{tail} to shift a stream by one word and \ttt{paste} to fuse (zip) two streams together.
\ttt{spell} is based on 
spell program~\cite{bentley1985spelling} 
The next pipeline is based on the original \ttt{spell} program developed by Johnson~\cite{bentley1985spelling}---another \unix classic:
  after some preprocessing, it makes clever use of \ttt{comm} to report words not in the dictionary.

Generally, the planning

@@
